Limited plasticity but increased variance in physiological rates across ectotherm populations under climate change

### 2025-02-17 ms.qmd
Dealt with all reviewer responses and responded to them all.

Effect sizes can be calculated from samples of organisms measured acutely at two temperatures or after having been acclimated these same temperatures ([@fig-concept]A). For studies that measure acute and acclimated responses we used the mean, standard deviation, and sample size to derive both acute and acclimation $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ estimates. For studies that only measured $R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$ after acclimation we could only calculate acclimation versions of these effect size estimates. For all effect sizes the higher temperature was in the numerator and the lower of the two temperatures in the denominator. As such, positive effect sizes indicate that the mean (i.e., $lnRR_{Q_{10}}$) or variance ($lnVR_{Q_{10}}$) is larger at the higher of the two temperatures (numerator) when standardized to 10°C. When measuring plasticity, it is the difference between $lnRR_{Q_{10}}$ acute (denoted, $lnRR_{Q_{10_{acute}}}$) and acclimation (denoted, $lnRR_{Q_{10_{acclimation}}}$) that captures the degree to which organisms plastically adjust (or acclimate). As done by @Seebacher2015-wi, we consider acute $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ as animals measured acutely at both temperatures even though one of the acute measurements is also the acclimation temperature. A better measure of "acute" responses would be to calculate $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ on two completely new temperatures but this was not often done in studies. Importantly, our effect sizes, as with $Q_{10}$ more generally, all assume that the effect of temperature on physiological rates (or changes in variance) is log-linear (see [@fig-concept]B & *Supplementary Materials* for further discussion). We test and control for any violations of these assumptions in our analysis (see below).

@@

  saveRDS(model6_bayes, file = "./output/models/lnVRQ10_6_bayes")

@@

Measurements of acute and acclimation responses from wild ectotherms were much less common than from captive populations ($N_{species}$ = `r length(unique(all_pred2_vr$spp))`, from `r length(unique(all_pred2_vr$uniques))` wild populations). Globally, there was a clear bias towards species in the Northern Hemisphere ([@fig-fig6]A-C). Projected changes in physiological rate variance were highly variable across the globe, however, variance was predicted to increase at all locations.

>>>>

Effect sizes can be calculated from samples of organisms measured acutely at two temperatures or after having been acclimated these same temperatures ([@fig-concept]A). For studies that measure acute and acclimated responses we used the mean, standard deviation, and sample size to derive both acute and acclimation $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ estimates. For studies that only measured $R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$ after acclimation we could only calculate acclimation versions of these effect size estimates. Ideally, all studies would have a fully factorial design but this was not always the case making it challenging to compare acute and acclimated responses within studies. Nonetheless, our analytical models are suitable for dealing with missing within-study acute effects (see below). In addition, analysis of a subset of the data where acute and acclimation effects could be compared within studies yields the same conclusion (See *Supplementary Materials*). For all effect sizes the higher temperature was in the numerator and the lower of the two temperatures in the denominator. As such, positive effect sizes indicate that the mean (i.e., $lnRR_{Q_{10}}$) or variance ($lnVR_{Q_{10}}$) is larger at the higher of the two temperatures (numerator) when standardized to 10°C. When measuring plasticity, it is the difference between $lnRR_{Q_{10}}$ acute (denoted, $lnRR_{Q_{10_{acute}}}$) and acclimation (denoted, $lnRR_{Q_{10_{acclimation}}}$) that captures the degree to which organisms plastically adjust (or acclimate). As done by @Seebacher2015-wi, we consider acute $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ as animals measured acutely at both temperatures even though one of the acute measurements is also the acclimation temperature. A better measure of "acute" responses would be to calculate $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ on two completely new temperatures but this was not often done in studies. Importantly, our effect sizes, as with $Q_{10}$ more generally, all assume that the effect of temperature on physiological rates (or changes in variance) is log-linear (see [@fig-concept]B & *Supplementary Materials* for further discussion). We test and control for any violations of these assumptions in our analysis (see below).

@@

# Model only used for model comparison
model6_bayes_nonsmooth <- brms::brm(effect_size | se(sqrt(V)) ~  max_temp_c + c_accl_time + type*habitat + Longitude + Latitude +
                               (1 + type | record_num) + 
                               (1 | gr(species_full, cov = A_wild_vr)) + 
                               (1 + max_temp_c | spp) + 
                               (1 | trait) + 
                               (1 | obs), 
                      data2 = list(A_wild_vr=A_wild_vr), family = gaussian(), 
                      prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                      control = list(adapt_delta = 0.99, max_treedepth = 12), 
                      chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, 
                      data = lnVRQ10_data_wild[complete.cases(lnVRQ10_data_wild$c_accl_time),])

  saveRDS(model6_bayes, file = "./output/models/lnVRQ10_6_bayes")
  saveRDS(model6_bayes_nonsmooth, file = "./output/models/lnVRQ10_6_bayes_nonsmooth")

@@

Measurements of acute and acclimation responses from wild ectotherms were much less common than from captive populations ($N_{species}$ = `r length(unique(all_pred2_vr$spp))`, from `r length(unique(all_pred2_vr$uniques))` wild populations). Globally, there was a clear bias towards species in the Northern Hemisphere ([@fig-fig6]A-C). Projected changes in physiological rate variance were highly variable across the globe, however, variance was predicted to increase at all locations. Latitude and longitude explained variation in these responses with models containing smoothers being supported over models with just main effects of latitude and longitude ($\Delat$ wAIC = 2.9).
